The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Work on section 4.3 and 4.4
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1808 4525493e-7705-40b1-a816-d608a930855b

\subsection{DAS core}
The DAS core library was responsible for data retrieval on demand. 
It had the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server was done via the REST \cite{REST} interface. 
%For each GET request we performed the following action.
Each input query was recorded into Analytics DB for further 
analysis. It was decomposed into the set of conditions and/or selection keys.
First the results were looked up in a cache and delivered to the requestor
if superset of input conditions was found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, was constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and store results into the cache. Finally, the results were
retrieved from the cache and delivered to the user.

The mapping between user queries, the DAS keys used in a query, 
and API names/parameters was stored in a Mapping DB. Such mapping was 
used by data-service plug-ins. For example
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
represents the mapping between \verb+site+ DAS keyword and
\verb+CMSNametoSE+ API provided by SiteDB data-service (the data-service
URLs were kept in the configuration file).
The data-service plug-in implementation were similar and their logic was 
abstracted, while a few extentions were required, e.g. specific 
security modules for data-services accessible via private network.

\subsection{DAS Mapping DB}
The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values, as regular expression patterns, into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used a \verb+logical_file_name+ notation to identify the name of the
logical file, while the same entity was called \verb+lfn+ 
in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key \verb+file+.
%\footnote{We
%also deal with physical file names, but still used {\it file} to refer to them in
%DAS notations, since simple regular expressions were used to identify the
%meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type \verb+file+ in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. In aforementioned example, 
usage of \verb+file+ in DAS query cause a call of DBS 
\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
This approach gave ability to end-users to express the meaning of their queries, especially
when they deal with numbers, and achieve accuracy of results. 
For instance, users were able to type \verb+run > 10 and run < 100+
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full keyword search is under development.
% within MongoDB.

\subsection{DAS Analytics DB}
The DAS Analytics DB played a special role. It collected information
about user requests placed to the system. Each request was recorded. Upon its
decomposition into set of selection keys and conditions we also recorded
API name and input parameters into the Analytics database and 
keep updating their counter information for repeated calls. 
This allows to keep track of frequency of API calls and establish 
pre-fetch strategies for most common queries.

@@

\subsection{DAS queries}
Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on analysis of daily work operations of our users, see \cite{DBS07}.
Each key had a set of attributes with primary one assigned to the key itself.
For instance, the \verb+block+ key had the following attributes
\verb+name+, \verb+size+, \verb+replica+ etc., 
with primary attribute being \verb+name+.
The choice of attributes were based on a set of meta-data information
associated with a given DAS key. Therefore end-users where able to specify
a context in their queries by using appropriate attribute(s). For example,
for query \verb+block+ we didn't perform a string matching to the
``block'' string, but rather look-up a block objects in a DAS, while query
\verb+block.name=abc*+ instructs system to look-up appropriate blocks
with matched names. At the same time, to complement the strength of the 
DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
search later.

>>>>

\subsection{DAS core\label{DAS_core}}
The DAS core library was responsible for data retrieval on demand. 
It had the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server was done via the REST \cite{REST} interface. 
%For each GET request we performed the following action.
Each input query was recorded into Analytics DB for further 
analysis. It was decomposed into the set of conditions and/or selection keys.
First the results were looked up in a cache and delivered to the requestor
if superset of input conditions was found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, was constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and stored results into the cache. At this step we also
performed necessary data aggregation for records with similar key-value
pairs. The data-service plug-in implementation were similar and their 
logic was abstracted, while a few extentions were required, e.g. specific 
security modules for data-services accessible via private network.
Finally, the results were retrieved from the cache and delivered to the user.

\subsection{DAS Mapping DB}
The Mapping DB represents DAS {\it registration} service. It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

%The Mapping DB hold the mapping between data-service APIs and DAS keys exposed to
%the users. It created a necessary glue between data-services and identifies
%how their data can be related to each other. Basically Mapping DB was positioned 
%as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which participate in DAS 
we stored their names, input parameters and associated DAS keys
into the Mapping DB. For example
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
represents DAS mapping record.\footnote{
The data-service URLs were kept in the configuration file.
}
This record can be read as following:
the \verb+CMSNametoSE+ API provided by 
\verb+sitedb+ data-service accepts \verb+name+ as input parameter. It
confirms regular expression pattern, \verb+re.compile('^T[0-3]_')+, and
corresponds to \verb+site+ DAS key. The \verb+site+ key
maps into \verb+site.name+ within DAS record.

Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
Each key was mapped into set of APIs whose output was aggregated afterwards.

\subsection{DAS Analytics DB}
The DAS Analytics DB was used to collect information
about user requests placed to the system. Each request was recorded. Upon its
decomposition into set of selection keys and conditions we also recorded
API name and input parameters into the Analytics DB and 
kept updating their counter information for repeated calls. 
This allowed to keep track of frequency of API calls and establish 
pre-fetch strategies for most common queries.

@@

\subsection{DAS queries}
Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on analysis of daily work operations of our users, see \cite{DBS07}.
Each key had a set of attributes with primary one assigned to the key itself.
For instance, the \verb+block+ key had the following attributes
\verb+name+, \verb+size+, \verb+replica+ etc., 
with primary attribute being \verb+name+.
The choice of attributes were based on a set of meta-data information
associated with a given DAS key. Therefore end-users where able to specify
a context in their queries by using appropriate attribute(s). For example,
for query \verb+block+ we didn't perform a string matching to the
``block'' string, but rather look-up a block objects in a DAS, while query
\verb+block.name=abc*+ instructs system to look-up appropriate blocks
with matched names. The numerical values, conditions, groupping were
easy to express too. For example, \verb+run > 10 and run < 100+ query 
was used to get all run records for provided range of run numbers. 
At the same time, to complement the strength of the 
DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
search later.
